---
title: "Evolution of Science"
author: "Rodrigo Valdes Ortiz"
date: "5/22/2018"
output: flexdashboard::flex_dashboard
# output: html_document
runtime: shiny
---

```{r}
library(data.table)
library(plyr)
library(ggplot2)
# setwd("/Users/ruy/Documents/UChicago/spring2018/data_viz/a2")
```

<style type="text/css">
body{ /* Normal  */
      font-size: 18px;
  }
</style>


```{r}
# Load the data
disciplines = fread("data.csv")
summary = count(disciplines, c('year','discipline'))
```



Graphs
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
# shiny inputs defined here
selectInput("dis", label = "Discipline:",
            choices = c("Literature", "Biology", "ComputerScience", "Economics", "PolScience", "Medicine", "Physics"), selected = "Literature")
```

```{r echo=FALSE}
# Main theme
blank_theme <- theme_minimal()+
  theme(
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )
```


Column
-------------------------------------

\newline

**Number of papers by academic field by year**

\newline

```{r}
renderPlot({
  plot_data = summary[summary$discipline == input$dis,]

  ggplot(plot_data, aes(x=year))  + 
    geom_line(aes(y=freq),color="indianred4", size = 5) +
    labs(title = input$dis, y = "Papers in the cluster", x = "Year") +
    scale_x_continuous(breaks=c(1985, 1990, 1995, 2000)) +
    scale_y_continuous(breaks=seq(0,550, by = 100)) +
    blank_theme +
    theme(legend.position="none") + 
    theme(axis.text.x = element_text(colour="grey20",size=14,face="plain"),
          axis.text.y = element_text(colour="grey20",size=14,face="plain"),
          axis.title.x = element_text(colour="grey20",size=15,face="bold"),
          axis.title.y = element_text(colour="grey20",size=15,face="bold"),
          plot.title = element_text(hjust = 0.5, size = 20))
})

```

**Percent change in the number of papers by academic field**

\newline

```{r}

renderPlot({
  
  plot_data = summary[summary$discipline == input$dis,]
  diference = diff(plot_data$freq, differences = 1)
  diference[16] = diference[15]
  plot_data$diference = diference
  plot_data$change = (plot_data$diference / plot_data$freq) * 100
  plot_data$zero = 0
  
  name_t = paste("Percent change in ", input$dis, sep=" ")

  ggplot(plot_data, aes(x=year))  + 
    geom_line(aes(y=change), color="navyblue", size = 5) +
    geom_line(aes(y=zero), color="red", size = 1) +
    labs(title = name_t , y = "Percent change", x = "Year") +
    scale_x_continuous(breaks=c(1985, 1990, 1995, 2000)) +
    scale_y_continuous(breaks=seq(-50, 50, by = 10)) +
    blank_theme +
    theme(legend.position="none") + 
    theme(axis.text.x = element_text(colour="grey20",size=14,face="plain"),
          axis.text.y = element_text(colour="grey20",size=14,face="plain"),
          axis.title.x = element_text(colour="grey20",size=15,face="bold"),
          axis.title.y = element_text(colour="grey20",size=15,face="bold"),
          plot.title = element_text(hjust = 0.5, size = 20))
  
  
})


```


Explanation
=====================================  

***History***

This project about identifying diverse patterns in the evolution of academic disciplines from 1985 to 2000. I depict two subsets of information. First, data about the number of papers in each academic discipline, which gives an idea of the actual size of the field. Second, percent change in the number of papers by year, which provides an intuition on the trends according to a different time frame. 

One example of why I present both graphs is physics. Note that the first graph might give an illusion that the field has increased exponentially. However, the second graph shows that its main growth was at the beginning of the data, and then is a plateau.

***Why this graphical form?***

I want to show the evolution of academic disciplines through time. Given that I want to show long-term patterns, I believe that line plots are excellent for this. Lines gives a sense of continuity against bar plots.

***Role of interactivity***

The main role of the App is to show to non-technical savvy people trends by academic discipline. For instance, this is designed to analyze field by field. Comparisons among different fields are difficult to see here. 

Users foster their creativity by changing the different disciplines and observe its trends. Some of the questions that user can generate are: why the exponential growth in physics was during the eighties?, why literature has decreased so much?, are labels consistent through time?


***NOTE***

All this is **simulated data** due to propietary data issues. 



